---
title: "Spatial Sampling Scenarios for Tanzania"
author: "Ernest Guevarra"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load libraries
if (!require(remotes)) install.packages("remotes")
remotes::install_github("spatialworks/tanzania")
library(rgdal)
library(rgeos)
library(raster)
library(tanzania)

## Read tif of gridded livestock population
livestockPop <- raster("2_GlobalRuminantLPS_GIS/glps_gleam_61113_10km.tif")
lsPopStudyArea <- raster::intersect(livestockPop, studyArea)
```

## Context
A study to be conducted in the northeastern border of Tanzania from Arusha up to west to Serengeti that will collect human population samples/information, livestock samples/information and also land use/land disturbance information. The question is what possible sampling approaches can be used that can unify these different sources of data for the study.

```{r map1, echo = FALSE, fig.cap = "Northeastern border of Tanzania from Arusha to Serengeti", fig.align = "center"}
## Subset region to study area
studyArea <- region[region$Region_Nam %in% c("Arusha", "Mara"), ]

## Plot study area
par(mar = c(0,0,0,0))
plot(studyArea, lty = 0)
plot(region, lwd = 1, add = TRUE)
plot(studyArea, lwd = 2, col = "gray70", add = TRUE)

## Inset
par(plt = c(0.79, 0.99, 0.79, 0.99), new = TRUE)
plot(region, lwd = 0.5)
plot(studyArea, col = "gray50", add = TRUE)
polygon(x = c(bbox(studyArea)[1, 1] - 0.2, bbox(studyArea)[1, 1] - 0.2,
              bbox(studyArea)[1, 2] + 0.2, bbox(studyArea)[1, 2] + 0.2),
        y = c(bbox(studyArea)[2, 1] - 0.2, bbox(studyArea)[2, 2] + 0.2,
              bbox(studyArea)[2, 2] + 0.2, bbox(studyArea)[2, 1] - 0.2),
        border = "red", lwd = 1)
box(which = "plot", lty = "solid", lwd = 2)
```

Other factors/conditions to consider are:

1. Population density of humans decrease as population density of livestock increase and as population density of humans increase, population density of livestock decrease.

2. In order to study land disturbance properly, samples/information need to also be collected in areas where humand and livestock do not reside/stay.

Additional assumptions made:

1. Approach to study sampling is using cluster sampling (CS) rather than simple random sampling (SRS).

2. 

Given these initial parameters, we consider the following sampling options.

## 

```{r}
## Subset village to study areas
studyVillages <- village[village$REGNAME %in% c("Arusha", "Mara"), ]

par(mar = c(0,0,0,0))
plot(studyArea, lty = 0)
plot(lsPopStudyArea, add = TRUE)
plot(region, lwd = 1, add = TRUE)
plot(studyVillages, border = "darkgreen", add = TRUE)
```


