---
title: "Spatial Sampling Scenarios for Tanzania"
author: "Ernest Guevarra"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  echo = TRUE
)

## Load libraries
if (!require(remotes)) install.packages("remotes")
remotes::install_github("spatialworks/tanzania")
remotes::install_github("spatialworks/spatialsampler")

library(rgdal)
library(rgeos)
library(raster)
library(tanzania)
library(spatialsampler)
```

## Context
A study to be conducted in the northeastern border of Tanzania from Arusha up to west to Serengeti that will collect human population samples/information, livestock samples/information and also land use/land disturbance information. The question is what possible sampling approaches can be used that can unify these different sources of data for the study.

```{r map1, echo = FALSE, fig.cap = "Northeastern border of Tanzania from Arusha to Serengeti", fig.align = "center"}
## Subset region to study area
studyArea <- region[region$Region_Nam %in% c("Arusha", "Mara"), ]

## Plot study area
par(mar = c(0,0,0,0))
plot(studyArea, lty = 0)
plot(region, lwd = 1, add = TRUE)
plot(studyArea, lwd = 2, col = "gray70", add = TRUE)

## Inset
par(plt = c(0.79, 0.99, 0.79, 0.99), new = TRUE)
plot(region, lwd = 0.5)
plot(studyArea, col = "gray50", add = TRUE)
polygon(x = c(bbox(studyArea)[1, 1] - 0.2, bbox(studyArea)[1, 1] - 0.2,
              bbox(studyArea)[1, 2] + 0.2, bbox(studyArea)[1, 2] + 0.2),
        y = c(bbox(studyArea)[2, 1] - 0.2, bbox(studyArea)[2, 2] + 0.2,
              bbox(studyArea)[2, 2] + 0.2, bbox(studyArea)[2, 1] - 0.2),
        border = "red", lwd = 1)
box(which = "plot", lty = "solid", lwd = 2)
```

Other factors/conditions to consider are:

1. Population density of humans decrease as population density of livestock increase and as population density of humans increase, population density of livestock decrease.

2. In order to study land disturbance properly, samples/information need to also be collected in areas where humand and livestock do not reside/stay.

Additional assumptions made:

1. Approach to study sampling is using cluster sampling (CS) rather than simple random sampling (SRS).

2. Study will use the data perform general estimation of specific indicators but mostly to use the data and the indicators for modelling to address study questions.

This document provides some ideas/insights as to how sampling for a study can be done in the given context and the stated parameters and assumptions.

## Data used for sampling scenario building

To build the sampling scenarios, the following datasets were retrieved and utilised:

1.

2. 




## Sampling scenarios

### Cluster sample selected proportional to population size

### Spatial sampling

```{r basic_spatial, echo = TRUE, fig.align = "center"}
## Subset village to study areas
studyVillages <- village[village$REGNAME %in% c("Arusha", "Mara"), ]

par(mar = c(0,0,0,0))

## Plot base maps
plot(studyArea, lty = 0)
plot(studyVillages, border = "darkgreen", add = TRUE)
plot(studyArea, lwd = 2, add = TRUE)

## Create sampling grid (n = 30 square grids)
studySP <- create_sp_grid(x = studyArea, 
                          n = 30,
                          country = "Tanzania", 
                          buffer = 2, type = "csas")

## Get nearest point
studyVillagesSP <- data.frame(coordinates(studyVillages), studyVillages@data)
names(studyVillagesSP) <- c("x", "y", names(studyVillages@data))

studySampleVillages <- get_nearest_point(data = studyVillagesSP,
                                         data.x = "x", data.y = "y",
                                         query = studySP, n = 1)

plot(studySP, cex = 0.3, add = TRUE)
points(studySampleVillages[ , c("x", "y")], cex = 0.5, pch = 0, col = "blue")
```

```{r sample-table, echo = FALSE}
tab <- studySampleVillages[ , c("spid", "REGNAME", "DISTNAME", "WARD_NAME", "STREET", "TOTAL")]
names(tab) <- c("SPID", "Region", "District", "Ward", "Street", "Population")

knitr::kable(x = tab)
```


```{r livelihood_zones, echo = FALSE, eval = FALSE, fig.align = "center"}
## Subset village to study areas
studyVillages <- village[village$REGNAME %in% c("Arusha", "Mara"), ]

plot(studyArea, lty = 0)
plot(livelihood_zone, lwd = 2, border = "red", add = TRUE)
plot(region, lwd = 2, border = "blue", add = TRUE)
#plot(studyVillages, border = "darkgreen", add = TRUE)
```


