---
title: '`r paste("Spatial Sampling Plan for ", params$country, " Study", sep = "")`'
date: '`r params$ts`'
output: html_document
params:
  ts: NA
  country: NA
  nSamplingUnits: NA
  buffer: NA
  samplingFrame: NA
  studyArea: NA
  samplingPoints: NA
  samplingGrid: NA
  adminArea: NA
  baseLayer: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  echo = FALSE
)
```

<br/>   
<br/>    

This document is a summary report of the spatial sampling plan for the study in `r params$country`.
    
<br/>   
<br/>

## Spatial sampling parameters

The spatial sampling plan was based on the following sampling parameters:

```{r parameters}
parameters <- c("Study country",
                "Number of sampling units",
                "Spatial sampling buffer")

parameterValues <- c(params$country, params$nSamplingUnits, params$buffer)

parameters <- data.frame(parameters, parameterValues)

knitr::kable(x = parameters) %>%
  kableExtra::kable_styling(full_width = FALSE)
```
    
<br/>   
<br/>
    
## Sampling frame

Based on the abovementioned parameters, the following sampling frame was arrived at:

```{r frame}
x <- params$samplingFrame

knitr::kable(x = x) %>%
  kableExtra::kable_styling(full_width = TRUE)
```
    
<br/>
    
The sampling frame can be downloaded as an `XLSX` file [here](`r paste("https://aegypti.ecohealthalliance.org/shiny/guevarra/spatialsampling/reports/", tolower(params$country), "_", params$ts, ".xlsx", sep = "")`).
    
<br/>
    
```{r map, fig.align = "center", fig.width = 10, fig.height = 10}
plot(params$studyArea, lty = 0)
plot(params$adminArea, lwd = 1, add = TRUE)
plot(params$studyArea, lwd = 3, 
     border = "darkgreen", 
     col = rgb(red = col2rgb("gray70", alpha = 0.7)[1],
               green = col2rgb("gray70", alpha = 0.7)[2],
               blue = col2rgb("gray70", alpha = 0.7)[3],
               alpha = 100,
               maxColorValue = 255),
     add = TRUE)
plot(params$samplingGrid, lwd = 2, border = "blue", add = TRUE)
points(params$samplingPoints@coords, pch = 20, col = "red", cex = 1)
```
    
<br/>
    
```{r leafletMap, out.width = "100%", out.height = "600px"}
leaflet() %>%
  addMapboxTiles(style_id = params$baseLayer,
                 username = "ernestguevarra") %>%
  setView(lng = coordinates(params$studyArea)[1], 
          lat = coordinates(params$studyArea)[2], 
          zoom = 9) %>%
  addPolygons(data = params$adminArea,
    color = "yellow",
    fill = FALSE,
    weight = 5,
    group = "Administrative boundaries") %>%
  addPolygons(data = params$studyArea,
    color = "green",
    fill = FALSE,
    weight = 10,
    group = "Study area") %>%
  addPolygons(data = params$samplingGrid,
    color = "blue",
    fill = FALSE,
    weight = 5,
    group = "Sampling grid") %>%
  addCircleMarkers(
    lng = params$samplingPoints@coords[ , 1],
    lat = params$samplingPoints@coords[ , 2],
    color = "red",
    radius = 5,
    weight = 2,
    group = "Sampling points")
```

